---
title: "Circular Dichroism"
author: "Christopher E. Berndsen"
date: "February 23, 2018"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(shiny)
```
## Protein Secondary Structure

Secondary structure of proteins are interactions between non-consecutive amino acids within a single polypeptide or protein chain. Secondary structures fall into three main categories:

1. Helices

2. Beta strands

3. Random Coils


There are sub-categories within each of these three structures, however we will limit our discussion to these three for now. **Helices** are formed when amino acids form a hydrogen bond between the backbone carbonyl of one amino acid and the backbone amine of an amino acid 3 to 4 positions away. **Beta Strands** are a fully extended chain of amino acids, which typically form hydrogen bonds with other beta strands to form **beta sheets**. When a structure falls into neither of these categories we usually say the region is unstructured or in a **random coil**. 

More information on [secondary structure](https://www.ncbi.nlm.nih.gov/books/NBK22580/) from NCBI. 


## Circular Dichroism

Most amino acids are **chiral** and therefore proteins are also chiral. The chirality of proteins means that they will preferentially absorb polarized light in one direction. The unequal absorbance of circularly polarized light is called **circular dichroism (CD)** and measuring the CD of proteins is the standard method for determining secondary structure. For more information and explanantions on CD spectroscopy of proteins see [Using circular dichroism spectra to estimate protein secondary structure](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2728378/) by Norma Greenfield. 

The different structures of protein absorb circularly polarized light and therefore have distinct CD spectra. 

```{r CD spectra, echo=FALSE, warning=FALSE, message=FALSE, fig.align="center"}
#Dataset for displaying
CDdat <- data.frame(lambda = seq(190, 250, by = 0.1))

CDdat <- CDdat %>% mutate(helix = 1*10^8 * (2230060.04151075*lambda^0 +
                                             -100548.516559741*lambda^1 +
                                             2037.18080475746*lambda^2 + 
                                             -24.4244919907991*lambda^3 +
                                             0.19190243015954*lambda^4 +
                                             -0.00103245782924168*lambda^5 +
                                             0.00000385211889091252*lambda^6 +
                                             -9.84175959744622E-09*lambda^7 +
                                             1.64786777298595E-11*lambda^8 +
                                             -1.63282751503442E-14*lambda^9 +
                                             7.27089674019501E-18*lambda^10)) %>% 
                  mutate(beta = 1*10^8 * (-677807.330017282*lambda^0 +
                                            30975.2707887604*lambda^1 +
                                            -636.143263740698*lambda^2 + 
                                            7.73164864362657*lambda^3 +
                                            -6.15861633716145E-02*lambda^4 +
                                            3.35943314432255E-04*lambda^5 +
                                            -1.27092416556044E-06*lambda^6 +
                                            3.29272089581372E-09*lambda^7 +
                                            -5.59118151750062E-12*lambda^8 +
                                            5.61899389095424E-15*lambda^9 +
                                            -2.53794637234403E-18*lambda^10)) %>%
                  mutate(coil = 1*10^8 * (-580939.072386969*lambda^0 +
                                            25845.2673351998*lambda^1 +
                                            -516.713088253122*lambda^2 + 
                                            6.1134023680003*lambda^3 +
                                            -4.74021175198809E-02*lambda^4 +
                                            2.51692531821056E-04*lambda^5 +
                                            -9.26824208397782E-07*lambda^6 +
                                            2.33714935193268E-09*lambda^7 +
                                            -3.86247107852678E-12*lambda^8 +
                                            3.77764956561175E-15*lambda^9 +
                                            -1.6603998403172E-18*lambda^10))



#Generate plots for pure secondary structures
ggplot() +
    geom_point(data = CDdat, aes(x = lambda, y = helix), color = "red") +
    geom_point(data = CDdat, aes(x = lambda, y = beta), color = "blue") +
    geom_point(data = CDdat, aes(x = lambda, y = coil), color = "magenta") +
    scale_x_continuous(breaks = seq(190, 250, by = 5)) +
    labs(x = "wavelength (nm)", y = "Ellipticity", title = "CD spectrum of protein seconday structures") +
    theme(axis.text = element_text(size = 10, face = "bold"), axis.title = element_text(size = 16, face = "bold")) +
    annotate("text", x = 202, y = 50000, label = "Helical", color = "red", size = 5) +
    annotate("text", x = 195, y = 20000, label = "Beta \nSheet", color = "blue", size = 5) +
    annotate("text", x = 195, y = -22000, label = "Random \nCoil", color = "magenta", size = 5)




```


The data in the figure above shows that helical structures have two minima at 222 nm and ~212 nm, while the spectrum for beta sheets will have a minimum at 217 nm. Proteins that are in random coil structures will have a minimum around 195 nm. 


## CD in monitoring protein structure changes

In addition to measuring protein secondary structures, CD can be used to monitor how protein structure is changing in response to environmental conditions such as salts ([example](https://www.sciencedirect.com/science/article/pii/S0301462213001361?via%3Dihub)) or temperature ([example](https://pubs.acs.org/doi/10.1021/acs.biochem.5b01362)). Whether the overall structure is changing or if the protein is unfolding, both will likely show changes in protein secondary structure. 

The data below show how the CD spectrum for an alpha helical would proceed as the protein unfolds or "melts" in response to increasing temperature.

```{r helix Melting curve, echo=FALSE, warning=FALSE, message=FALSE, fig.align="center"}

#Make melting curve
CDmelt <- CDdat %>% mutate('20C' = 1*helix + 0*beta + 0*coil) %>%
  mutate('30C' = 0.95*helix + 0*beta + 0.05*coil) %>%
  mutate('40C' = 0.85*helix + 0*beta + 0.2*coil) %>%
  mutate('50C' = 0.65*helix + 0*beta + 0.35*coil) %>%
  mutate('60C' = 0.4*helix + 0*beta + 0.6*coil) %>%
  mutate('70C' = 0.2*helix + 0*beta + 0.8*coil) %>%
  mutate('80C' = 0.05*helix + 0*beta + 0.95*coil) %>%
  mutate('90C' = 0*helix + 0*beta + 1*coil) %>%
  select(1, 5:12) %>%
  gather(Temperature, CD, 2:9)

ggplot(CDmelt, aes(x = lambda, y = CD, color = Temperature)) +
  geom_point(alpha = 0.4) +
  scale_x_continuous(breaks = seq(190, 250, by = 5)) +
  labs(x = "wavelength (nm)", y = "Ellipticity", title = "Melting of a helical protein") +
  theme(axis.text = element_text(size = 10, face = "bold"), axis.title = element_text(size = 16, face = "bold"))






```


Below is a similar plot but for a beta sheet protein

```{r sheet Melting curve, echo=FALSE, warning=FALSE, message=FALSE, fig.align="center"}

#Make melting curve
CDmeltbeta <- CDdat %>% mutate('20C' = 0*helix + 1*beta + 0*coil) %>%
  mutate('30C' = 0*helix + 0.95*beta + 0.05*coil) %>%
  mutate('40C' = 0*helix + 0.85*beta + 0.2*coil) %>%
  mutate('50C' = 0*helix + 0.65*beta + 0.35*coil) %>%
  mutate('60C' = 0*helix + 0.4*beta + 0.6*coil) %>%
  mutate('70C' = 0*helix + 0.5*beta + 0.8*coil) %>%
  mutate('80C' = 0*helix + 0.05*beta + 0.95*coil) %>%
  mutate('90C' = 0*helix + 0*beta + 1*coil) %>%
  select(1, 5:12) %>%
  gather(Temperature, CD, 2:9)

#Plot data
ggplot(CDmeltbeta, aes(x = lambda, y = CD, color = Temperature)) +
  geom_point(alpha = 0.4) +
  scale_x_continuous(breaks = seq(190, 250, by = 5)) +
  labs(x = "wavelength (nm)", y = "Ellipticity", title = "Melting of a beta sheet protein") +
  theme(axis.text = element_text(size = 10, face = "bold"), axis.title = element_text(size = 16, face = "bold"))


```




```{r CD, echo=FALSE}
#input
inputPanel(
  selectInput("mode",
                   "Which mode do you want to use?",
                   choices = list("Prediction" = "predict",
                                  "Display" = "display")
        )
        ,
           
     mainPanel(
         #conditional panel 
         conditionalPanel(
           condition = "input.mode == 'predict'",
           checkboxInput("guides", 
                         "Show secondary structure guides?", 
                         value = FALSE, 
                         width = NULL)
           ,
           numericInput("helix",
                        "% alpha helix",
                        min = 0,
                        max = 100,
                        step = 5,
                        value = 50)
           ,
           numericInput("sheet",
                        "% beta sheet",
                        min = 0,
                        max = 100,
                        step = 5,
                        value = 50)
           ,
           numericInput("coil",
                        "% random coil",
                        min = 0,
                        max = 100,
                        step = 5,
                        value = 0)
           )
         ,
         conditionalPanel(
           condition = "input.mode == 'display'",
           selectInput("protein",
                       "Pick a protein to show predicted data for",
                       choices = list("Lysozyme (1LYD)" = "lyso",
                                      "Ubiquitin (1UBQ)" = "ub",
                                      "BST2 (3MQB)" = "bst",
                                      "Hemoglobin (2HHB)" = "hemo",
                                      "Antibody (1IGT)" = "ab")
           )
           ,
           "Simulated numbers generated in YASARA using the PDB IDs indicated"
           )
         )
)

#output
renderPlot({
       if(input$mode == "predict") {
       #check to make sure fractions add to 100%          
       validate(
         need(input$helix/100 + input$sheet/100 + input$coil/100 == 1, 
              "% of helix, sheet, and coil must = 100%"
              )
       )
     
     
       #generate basis set
       CDdat <- data.frame(lambda = seq(190, 250, by = 0.2))
       CDdat <- CDdat %>% mutate(helix = 1*10^8 * (2230060.04151075*lambda^0 +
                                                   -100548.516559741*lambda^1 +
                                                   2037.18080475746*lambda^2 + 
                                                   -24.4244919907991*lambda^3 +
                                                   0.19190243015954*lambda^4 +
                                                   -0.00103245782924168*lambda^5 +
                                                   0.00000385211889091252*lambda^6 +
                                                   -9.84175959744622E-09*lambda^7 +
                                                   1.64786777298595E-11*lambda^8 +
                                                   -1.63282751503442E-14*lambda^9 +
                                                   7.27089674019501E-18*lambda^10)) %>% 
       mutate(beta = 1*10^8 * (-677807.330017282*lambda^0 +
                                 30975.2707887604*lambda^1 +
                                 -636.143263740698*lambda^2 + 
                                 7.73164864362657*lambda^3 +
                                 -6.15861633716145E-02*lambda^4 +
                                 3.35943314432255E-04*lambda^5 +
                                 -1.27092416556044E-06*lambda^6 +
                                 3.29272089581372E-09*lambda^7 +
                                 -5.59118151750062E-12*lambda^8 +
                                 5.61899389095424E-15*lambda^9 +
                                 -2.53794637234403E-18*lambda^10)) %>%
       mutate(coil = 1*10^8 * (-580939.072386969*lambda^0 +
                                 25845.2673351998*lambda^1 +
                                 -516.713088253122*lambda^2 + 
                                 6.1134023680003*lambda^3 +
                                 -4.74021175198809E-02*lambda^4 +
                                 2.51692531821056E-04*lambda^5 +
                                 -9.26824208397782E-07*lambda^6 +
                                 2.33714935193268E-09*lambda^7 +
                                 -3.86247107852678E-12*lambda^8 +
                                 3.77764956561175E-15*lambda^9 +
                                 -1.6603998403172E-18*lambda^10))
     
       #Predict spectrum based on user input  
       CDdat <- CDdat %>% mutate(prediction = (input$helix)/100*helix + (input$sheet/100)*beta + (input$coil/100)*coil)
      
      # draw the plot
       if(input$guides == FALSE) {
         ggplot(CDdat, aes(x = lambda, y = prediction/1000), color = "red") +
           geom_jitter(alpha = 0.4) +
           scale_x_continuous(breaks = seq(190, 250, by = 5)) +
           labs(x = "wavelength (nm)", y = "Ellipticity", title = "Predicted CD spectrum") +
           geom_hline(yintercept = 0) +
           ylim(-50, 75) +
           theme_classic() +
           theme(axis.text = element_text(size = 10, face = "bold"), axis.title = element_text(size = 16, face = "bold"))
       }
       else {
         ggplot() +
           geom_jitter(data = CDdat, aes(x = lambda, y = prediction/1000), fill = "red", alpha = 0.4) +
           geom_line(data = CDdat, aes(x = lambda, y = helix/1000), color = "red") +
           geom_line(data = CDdat, aes(x = lambda, y = beta/1000), color = "green") +
           geom_line(data = CDdat, aes(x = lambda, y = coil/1000), color = "purple") +
           geom_hline(yintercept = 0) +
           scale_x_continuous(breaks = seq(190, 250, by = 5)) +
           labs(x = "wavelength (nm)", y = "Ellipticity", title = "Predicted CD spectrum") +
           annotate("text", x = 235, y = 25, label = "Helix", color = "red", size = 5) +
           annotate("text", x = 235, y = 20, label = "Beta Sheet", color = "green", size = 5) +
           annotate("text", x = 235, y = 15, label = "Random Coil", color = "purple", size = 5) +
           ylim(-50, 75) +
           theme_classic() +
           theme(axis.text = element_text(size = 10, face = "bold"), axis.title = element_text(size = 16, face = "bold"))
        
       }
       }
     else {
       #Generate the wavelength values
       CDdat <- data.frame(lambda = seq(190, 250, by = 0.2))
       
       #Generate the basis set from Abriata, L., J. Chem. Educ., 2011, 88 (9), pp 1268–1273 and Davidson, B. and Fasman, G. D., Biochemistry 1967 6 (6) 1616-1629
       CDdat <- CDdat %>% mutate(helix = 1*10^8 * (2230060.04151075*lambda^0 +
                                                     -100548.516559741*lambda^1 +
                                                     2037.18080475746*lambda^2 + 
                                                     -24.4244919907991*lambda^3 +
                                                     0.19190243015954*lambda^4 +
                                                     -0.00103245782924168*lambda^5 +
                                                     0.00000385211889091252*lambda^6 +
                                                     -9.84175959744622E-09*lambda^7 +
                                                     1.64786777298595E-11*lambda^8 +
                                                     -1.63282751503442E-14*lambda^9 +
                                                     7.27089674019501E-18*lambda^10)) %>% 
         mutate(beta = 1*10^8 * (-677807.330017282*lambda^0 +
                                   30975.2707887604*lambda^1 +
                                   -636.143263740698*lambda^2 + 
                                   7.73164864362657*lambda^3 +
                                   -6.15861633716145E-02*lambda^4 +
                                   3.35943314432255E-04*lambda^5 +
                                   -1.27092416556044E-06*lambda^6 +
                                   3.29272089581372E-09*lambda^7 +
                                   -5.59118151750062E-12*lambda^8 +
                                   5.61899389095424E-15*lambda^9 +
                                   -2.53794637234403E-18*lambda^10)) %>%
         mutate(coil = 1*10^8 * (-580939.072386969*lambda^0 +
                                   25845.2673351998*lambda^1 +
                                   -516.713088253122*lambda^2 + 
                                   6.1134023680003*lambda^3 +
                                   -4.74021175198809E-02*lambda^4 +
                                   2.51692531821056E-04*lambda^5 +
                                   -9.26824208397782E-07*lambda^6 +
                                   2.33714935193268E-09*lambda^7 +
                                   -3.86247107852678E-12*lambda^8 +
                                   3.77764956561175E-15*lambda^9 +
                                   -1.6603998403172E-18*lambda^10))
       
         #Predict spectrum based on user input  
         CDdat <- CDdat %>% 
         mutate(display = case_when(
           input$protein == 'lyso' ~ 0.7*helix + 0.07*beta + 0.17*coil,
           input$protein == 'ub' ~ 0.17*helix + 0.40*beta + 0.43*coil,
           input$protein == 'bst' ~ 0.92*helix + 0*beta + 0.08*coil,
           input$protein == 'hemo' ~ 0.74*helix + 0*beta + 0.17*coil,
           input$protein == 'ab' ~ 0.05*helix + 0.46*beta + 0.36*coil))
         
        ggplot() +
         geom_jitter(data = CDdat, aes(x = lambda, y = display/1000), fill = "red", alpha = 0.4) +
         geom_line(data = CDdat, aes(x = lambda, y = helix/1000), color = "red") +
         geom_line(data = CDdat, aes(x = lambda, y = beta/1000), color = "green") +
         geom_line(data = CDdat, aes(x = lambda, y = coil/1000), color = "purple") +
         geom_hline(yintercept = 0) +
         scale_x_continuous(breaks = seq(190, 250, by = 5)) +
         labs(x = "wavelength (nm)", y = "Ellipticity") +
         annotate("text", x = 235, y = 27, label = "Helix", color = "red", size = 7) +
         annotate("text", x = 235, y = 20, label = "Beta Sheet", color = "green", size = 7) +
         annotate("text", x = 235, y = 13, label = "Random Coil", color = "purple", size = 7) +
         ylim(-50, 75) +
         theme_classic() +
         theme(axis.text = element_text(size = 10, face = "bold"), axis.title = element_text(size = 16, face = "bold"))
         
         
         
     }
})
   
   
renderText({
     if(input$mode == "predict") {
       paste("Spectrum showing", input$helix, "% alpha helix", input$sheet, "% beta sheet", input$coil, "% random coil")
       
     }
     else{
     label <- case_when(
           input$protein == 'lyso' ~ "70% helix, 7% beta strand, 17% random coil",
           input$protein == 'ub' ~ "17% helix, 40% beta strand, 43% random coil",
           input$protein == 'bst' ~ "92% helix, 0% beta strand, 8% random coil",
           input$protein == 'hemo' ~ "74% helix, 0% beta strand, 17% random coil",
           input$protein == 'ab' ~ "5% helix, 46% beta strand, 36% random coil")
       
           paste("Structure contains", label)
     }
   })

```
