---
title: "Data Dictionary"
author: "Berndsen, Roy, and Sutton"
date: "Developed January 26, 2018"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(ggrepel)
library(reshape2)
library(readr)
library(readxl)
library(broom)
library(cowplot)
library(kableExtra)

```

### REVISED: `r Sys.Date()`

## analysis.tab

```{r message=FALSE, warning=FALSE,fig.height=5, fig.width=5}
#Load Data
data <- read_table("Clean_Tetherin_analysis.tab")

#rename data columns
colnames(data) <- c("Time[ps]", "Energy[kJ/mol]", "Bond", "Angle", "Dihedral", "Planarity", "Coulomb", "VdW", "RMSDs[A]:CA", "Backbone", "HeavyAtoms")

#Split table in two chunks for easier visualizing, do not do this during analysis
working <- data[1:6, 1:6]
workingtwo <- data[1:6, 7:11]

#Show data organization
knitr::kable(working, "html") %>% kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE)

knitr::kable(workingtwo, "html") %>% kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE)

```

### Table Definitions

**Time [ps]** = Time in picoseconds

**Energy [kJ/mol]** = Energy in kilojoules per mole at X time

**Bond** = Bond energy in kilojoules per mole at X time

**Angle** = Angle energy in kilojoules per mole at X time

**Dihedral** = Energy in dihedral angles in kilojoules per mole at X time

**Planarity** = Energy in planary bonds in kilojoules per mole at X time

**Coulomb** = Energy of Coulombic interactions in kilojoules per mole at X time

**VdW** = Energy of van der Waals in kilojoules per mole at X time

**RMSDs[A]:CA** = Root Mean Square Deviation from initial model in Angstroms for the Calpha atoms

**Backbone** = Root Mean Square Deviation from initial model in Angstroms for the backbone (Calpha, C, and N) atoms

**HeavyAtoms** = Root Mean Square Deviation from initial model in Angstroms for the all the non-H atoms including side chains



***

\pagebreak

## analysisres.tab

```{r message = FALSE, warning = FALSE}
#Load Data
datares <- read_table("Clean_Tetherin_analysisres.tab")
colnames(datares) <- c("Residue", "RMSDs[A]:CA", "Backbone", "HeavyAtoms")

#Clean Data
datares2 <- colsplit(datares$HeavyAtoms, " ", c("Heavy", "RMSF"))
datares <- bind_cols(datares, datares2)
datares <- datares %>% select(-HeavyAtoms)
colnames(datares) <- c("Residue", "RMSDs[A]:CA", "Backbone RMSD", "Heavy Atoms RMSD", "RMSF[A]")

#Show data organization
workingres <- datares[1:6, 1:5]
knitr::kable(workingres, "html") %>% kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE)

```

### Table Definitions

**Residue** = Amino acid in Tetherin in 3 letter code with position number and molecule assignment

**RMSDs[A]:CA** = Root Mean Square Deviation from initial model in Angstroms for the Calpha atoms

**Backbone RMSD** = Root Mean Square Deviation from initial model in Angstroms for the backbone (Calpha, C, and N) atoms

**Heavy Atoms RMSD** = Root Mean Square Deviation from initial model in Angstroms for the all the non-H atoms including side chains

**RMSF[A]** = Root Mean Square Flucuation for the back bone atoms in angstroms

***


\pagebreak

### dccm.tab


```{r message = FALSE, warning = FALSE}
#Load data
dccmdata <- read_table2("Clean_Tetherin_dccm.tab", col_names = FALSE, skip = 1)

#Create row and column names file
vec <- datares %>% filter(!is.na(Residue))
namesvec <- vec[['Residue']]

#name the columns in the dccm
colnames(dccmdata) <- c(namesvec)

#Add the row names
Residue <- datares %>% filter(!is.na(Residue)) %>% select(Residue)
dccmdata <- dccmdata[1:317, 1:317]
dccmdata <- bind_cols(Residue, dccmdata)

#Show the table
dccmdisplay <- dccmdata[1:7, 1:7]

knitr::kable(dccmdisplay, "html") %>% kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE)


```

### Table Definitions

**Residue** = The amino acid type, position number and molecule for the data.

**All other columns** = The normalized correlation of the movement of the amino acid in residue and the amijno acid in the column name. Range spans from -1 to +1 with -1 being anticorrelated and +1 full correlated. 


\pagebreak

### SecStrAnalysis.tab


```{r message = FALSE, warning = FALSE}
#Load required data sets
SecStr <- read_table2("Clean_Tetherin_SecStrAnalysis.tab", col_names = FALSE, skip = 2)
RMSf <- read_table("Clean_Tetherin_analysisres.tab")
colnames(RMSf) <- c("Residue", "RMSDs[A]:CA", "Backbone", "HeavyAtoms")

#split names column in RMSf
colnames(RMSf) <- c("Residue", "RMSDs[A]:CA", "Backbone", "HeavyAtoms")
RMSf <- RMSf %>% separate(., Residue, into = c("AAtype", "ResNum", "Mol"))

#Make final object and remove NA
RMSF <<- RMSf %>% filter(!is.na(ResNum))

#Create names vector
vec <- RMSF %>% select(1:3) %>% filter(!is.na(ResNum)) %>% unite(AA, AAtype, ResNum, Mol, sep = "_")
namesvec <- vec[['AA']]
  
#name the columns in the df
colnames(SecStr) <- c("Time", namesvec)

#Show part of the data
SecStrdisplay <- SecStr[1:7, 1:7]

knitr::kable(SecStrdisplay, "html") %>% kable_styling(bootstrap_options = c("striped", "condensed"), full_width = FALSE) 


```

### Table Definitions

**Time** = Time in picoseconds

**All other columns** = The assigned secondary structure for the amino acid indicated in the position. 
  
*Categories of secondary structure are:*

H = alpha helix (n to n+4 hydrogen bond)

E/EE = beta strand

T = turn

G = 3/10 helix (n to n+3 hydrogen bond)

I = pi helix (n to n+5 hydrogen bond)

C = everything not described above
